---
title: "DataMiningandHealthcare"
author: "Sophie Fournier"
date: '2022-08-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the Veteran's Administration Lung Cancer Trial data set (Kalbfeisch and Prentice).This data set shows the treatment used and the effectiveness dependent on a variety of factors.

There are 8 attributes:

stime: Survival time; days since treatment

status: 1(dead) and 0(alive)

treat: 1(standard treatment) and 2(test drug treatment)

age: in years

Karn: Karnofsky score

diag.time: time since diagnosis (months)

cell: type of cell; 1(squamous cell), 2(small cell), 3(adeno cell), 4(large cell)

prior: prior therapy (0=none and 10=yes) 

Decision Tree Mining is a probability tree that helps with decision making.

```{r}
# Import Data set
library(knitr)
library(rmarkdown)
library(MASS)
x <- VA
x
```

## Including Plots

You can also embed plots, for example:

```{r}
# Clean-up the Data set
library(dplyr)
clean_x <- x %>%
  mutate(status = factor(status, levels = c(1, 0), labels = c('Dead', 'Alive'))) %>%
  mutate(treat = factor(treat, levels = c(1, 2), labels = c('standard', 'test drug'))) %>% 
  mutate(cell = factor(cell, levels = c(1, 2, 3, 4), labels = c('squamous', 'small', 'adeno', 'large'))) %>%
  mutate(prior = factor(prior, levels = c(0, 10), labels = c('none', 'yes')))
x2 <- na.omit(clean_x)
x2
```

```{r}
# Split data set into train and test
# Common split is an 80/20 split 
# Create function
trainfunction <- function(data, size = 0.8, train = TRUE) {
   n_rows = nrow(data)
    totalrows = size * n_rows
    sample <- 1:totalrows
    if (train == TRUE) {
        return (data[sample, ])
    } else {
        return (data[-sample, ])
    }
}
traindata <- trainfunction(x2, size = 0.8, train = TRUE)
testdata <- trainfunction(x2, size = 0.8, train = FALSE)
traindata
```


```{r}
# Create Decision Tree
library(rpart)
library(rpart.plot)
VAtree <- rpart(stime~., data = traindata)
rpart.plot(VAtree)
```

```{r}
# Create scatter plot with box plots
library(ggplot2)
VAplot <- ggplot(traindata, aes(x=treat, y=age)) + geom_jitter(aes(size=status, color=status)) + geom_boxplot(aes(alpha = 0.3, outlier.shape = NA)) + labs(y="Age", x="Type of Treatment") + ggtitle("Effectiveness of Treatment based on Treatment Type and Age", subtitle = "Both Treatments were Ineffective for Most Patients")
VAplot
```